---
title: "R Notebook"
output: html_notebook
---
# Multiple Regression

```{r}
library(tidyverse)
library(skimr)
library(moderndive)
library(ISLR)
```

## One numerical and one categorical explanatory variable

### Interaction model

In the `evals` dataset, investigate the relationship between explanatory variables age (numerical) and gender (categorical), and score.

```{r}
evals_ch6 <- evals %>% 
  select(ID, age, gender, score)

glimpse(evals_ch6)
```
```{r}
evals_ch6 %>% sample_n(5)
```
```{r}
library(skimr)
evals_ch6 %>% select(age, gender, score) %>% skim()
```
```{r}
evals_ch6 %>% 
  summarise(r = cor(age, score))
```
```{r}
evals_ch6 %>% 
  ggplot(aes(x = age, y = score, color = gender)) +
    geom_point() +
    geom_smooth(method = 'lm', se = FALSE)
```

```{r}
library(broom)
score_model_interaction <- lm(score ~ age * gender, data = evals_ch6)
tidy(score_model_interaction)
```
In an interaction model like this, the reqgression lines
$$\hat{y} = \widehat{score} = b_0 + b_\textrm{age} \cdot \textrm{age} + b_\textrm{male} \cdot 1_\textrm{is male}(x) + b_\textrm{age,male} \cdot \textrm{age} \cdot 1_\textrm{is male}(x)$$
As was the case with the model using a single categorical explanatory variable, the indicator term $1_\textrm{is male}(x)$ zeroes out if the record is not male. The interaction effect, $b_\textrm{age,male}$ is `age:gendermale` from the regression table. This term indicates the extent to which one variable (`age`) depends on another (`gender`)

Thus the equation for the female line is:

$$\hat{y} = \widehat{score} = 4.883 - .018 \cdot age$$
Whereas for males it's:
$$
\hat{y} = \widehat{score} = (4.883 - .446) + (-.018 + 0.014) \cdot age \\
= 4.437 - 0.004 \cdot age
$$

### Parallel slopes

In an interaction model, the explanatory variables can behave differently across groups. In the `age * gender` model, the `age` penalty for men was significantly lower for men than for women.

In a parallel slopes model, only the intercept varies. Rather than `age * gender`, the model is `age + gender`.

```{r}
score_model_parallel_slopes <- lm(score ~ age + gender, data = evals_ch6)
tidy(score_model_parallel_slopes)
```

```{r}
score_model_parallel_slopes %>% 
  augment() %>% 
  ggplot(aes(x = age, y = score, color = gender)) +
    geom_point() +
    geom_line(aes(y = .fitted))
```

### Residuals

```{r}
score_model_interaction %>% 
  augment() %>% 
  select(score, age, gender, .fitted, .resid)
```

```{r}
score_model_parallel_slopes %>% 
  augment() %>% 
  select(score, age, gender, .fitted, .resid)
```
## Two Numerical Explanatory Variables

Model debt ($y$) as a factor of credit limit ($x_1$) and income ($x_2$).

```{r}
credit_ch6 <- Credit %>% 
  as_tibble() %>% 
    select(ID, debt = Balance, credit_limit = Limit, 
           income = Income, credit_rating = Rating, age = Age)

glimpse(credit_ch6)
```
```{r}
credit_ch6 %>% sample_n(6)
```
```{r}
credit_ch6 %>% select(debt, credit_limit, age) %>% skim()
```

```{r}
credit_ch6 %>% 
  summarise(r_debt_credit_limit = cor(debt, credit_limit),
            r_debt_age = cor(debt, age))
```
```{r}
credit_ch6 %>% 
  select(debt, credit_limit, age) %>% 
  cor()
```
```{r}
library(patchwork)
credit_ch6 %>%
  ggplot(aes(x = credit_limit, y = debt)) +
    geom_point() +
    geom_smooth(method = 'lm', se = FALSE) +
    xlab('credit limit in $') -> p1

credit_ch6 %>%
  ggplot(aes(x = income, y = debt)) +
    geom_point() +
    geom_smooth(method = 'lm', se = FALSE) +
    xlab('income in $1000') -> p2

p1 + p2
```
```{r}
debt_model <- lm(debt ~ credit_limit + income, data = credit_ch6)
tidy(debt_model)
```
This plots separately show the relationships between credit limit and debt and between income and debt, but in order to show the relationship between all 3, a 3D plot is needed, and the regression surface is a plane rather than a line. 

### Debt, Credit Rating and Age

```{r}
credit_ch6 %>% 
  select(debt, credit_rating, age) %>% 
  skim()
```
```{r}
credit_ch6 %>% 
  summarise(r_debt_credit_rating = cor(debt, credit_rating), r_debt_age = cor(debt, age))
```
```{r}
p1 <- credit_ch6 %>% 
  ggplot(aes(x = credit_rating, y = debt)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE)
p2 <- credit_ch6 %>% 
  ggplot(aes(x = age, y = debt)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE)

p1 + p2
```

### Regression Plane

```{r}
debt_model <- lm(debt ~ credit_limit + income, data = credit_ch6)
tidy(debt_model)
```
Even though credit limit and income are *separately* positively correlated with debt, they are jointly negatively correlated. This is Simpson's paradox, which states that adding an additional variable to a model can reverse or negate a trend observed before adding the new variable.

#### Regression Plane with Debt, Credit Rating and Age

```{r}
debt_credit_age_model <- lm(debt ~ credit_rating + age, data = credit_ch6)
tidy(debt_credit_age_model)
```